Dynamic Modeling, Parameter Estimation and Uncertainty Analysis in \proglang{R}

### 2016-10-03 article.tex
Abschnitt 3 fertig überarbeitet

In \pkg{dMod} different experimental conditions are handled by lists. Parameter transformations, prediction- and observation functions stacked by the \code{"+"}

@@

The aim of \pkg{dMod} is parameter estimation. The objective function to be minimized for this purpose is the weighted least squares function which is produced by the command \code{normL2(data, prdfn)} where \code{data} is a \code{datalist} object and \code{prdfn} is a prediction function. A \code{datalist} is a list of \code{data.frame}s providing time points, observables, measured values and uncertainties for each condition. The \code{as.datalist()} command allows to construct a datalist from a \code{data.frame} as it is typically used for mixed effects models or plotting with \pkg{ggplot2}. % Satz verwirrend

@@

Objective functions can be added by the \code{"+"} operator, meaning the objective values, gradients and Hessians are accumulated by summation. Besides \code{normL2()}, functions like \code{constraintL2()} and \code{datapointL2()} produce objective functions, too. They allow to define parameter priors or treat data points as parameters.

>>>>

In \pkg{dMod} different experimental conditions are handled by lists. Parameter transformations, prediction- and observation functions stacked by the \code{"+"} operator return list objects. On the other hand, \code{data.frame}s as they are used for linear modeling, mixed-effects modeling or plotting with \pkg{ggplot2} are highly convenient to organize the data. For the sake of computational efficiency, we keep the list structure of predictions and implement a class \code{datalist} to adapt the \code{data.frame} objects. A \code{datalist} is a list of data frames with the same structure, observable names, time points, measured values and measurement uncertainty.

Objects of class \code{datalist} are usually generated by the \code{as.datalist()} command from a \code{data.frame}. The factor variables in the data frame to be used as generators for the unique condition names can be passed by the \code{split.by} argument. The resulting list of data frames has an additional attribute \code{"condition.grid"}, a translation table between the condition names and the original factor variables which can be used for specification of parameter transformations or augmentation of predictions by descriptive columns.

@@

The aim of \pkg{dMod} is parameter estimation. The objective function to be minimized for this purpose is the weighted least squares function which is produced by the command \code{normL2(data, prdfn)} where \code{data} is a \code{datalist} object and \code{prdfn} is a prediction function.

@@

Objective functions can be added by the \code{"+"} operator, meaning the objective values, gradients and Hessians are accumulated by summation. Besides the standard function \code{normL2()}, the functions \code{constraintL2()} and \code{datapointL2()} produce objective functions, too. They allow to define parameter priors or treat data points as parameters. Thus, a typical objective function used for parameter estimation could be \code{obj <- normL2(data, g*x*(p1+p2+p3)) + constraintL2(mu, sigma)}.
